---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading required libraries

```{r  message=FALSE }
library(lubridate);library(dplyr);library(ggplot2); library(tidyr)
```

## Loading and preprocessing the data

```{r, echo = TRUE}
my_data <- read.csv("activity.csv")
my_data <- my_data %>% 
  mutate(date = ymd(date))
```

## What is mean total number of steps taken per day?

```{r, echo = TRUE}
sum_data <- my_data %>%
  group_by(day(date))%>%
  summarize(total_steps_perday = sum(steps)) 
```


```{r, echo = TRUE, message=FALSE, warning=FALSE}
# Make a histogram of the total number of steps taken each day and calculate and report the mean and median of the total number of steps taken per day
total_steps_perday_mean   = round(
                                mean(sum_data$total_steps_perday, na.rm=TRUE) 
                                )
total_steps_perday_median = round(
                                median(sum_data$total_steps_perday, na.rm=TRUE)
                                )
sum_data %>%
  ggplot(aes(x=total_steps_perday))+
  geom_histogram() + ylab("count") + xlab("Total steps per day")+ labs(title="Histogram of the total number of steps taken each day")+
  geom_vline(aes(xintercept = total_steps_perday_mean ),linetype=2,color="blue") + 
  annotate("text", x=total_steps_perday_mean-500, y=0.5, label=paste("mean=",total_steps_perday_mean), angle=90, size=5, color="blue",) +
  geom_vline(aes(xintercept = total_steps_perday_median ),linetype=2,color="red") + 
  annotate("text", x=total_steps_perday_median+500, y=0.5, label=paste("median=",total_steps_perday_median), angle=90, size=5, color="red") 
  
```

The mean is 21.1K steps while the median is 21.7K

## What is the average daily activity pattern?

Make a time series plot (i.e. type = "l"type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r, echo = TRUE, message=FALSE, warning=FALSE}
time_int_data = my_data %>%
  group_by(interval)%>%
  summarize(avr_steps_byinterval = mean(steps,na.rm=TRUE)) 

time_int_data_max <- arrange(time_int_data, desc(avr_steps_byinterval))[[1,1]]
time_int_data %>%
  ggplot(aes(x=interval , 
            y=avr_steps_byinterval)) + 
            geom_line() + xlab("5 minute intervals across 24 hours") + ylab("total steps per 5 min interval") + 
            labs(title="Time series plot of the average number of steps taken")+
            geom_vline(aes(xintercept = time_int_data_max),linetype=2, size=1, color="blue")  + 
            annotate("text", x=time_int_data_max+30, y=80, 
                     label=paste("interval with max avr steps)=", 
                                 time_int_data_max), angle=90, size=5, color="blue",)

```

Question:> Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
Answer:    at 835th interval which is around 13:55 is the interval with the highest no. of steps

## Imputing missing values
```{r, echo = TRUE ,message=FALSE, warning=FALSE}


data <- read.csv("activity.csv")
tidied_data <- data %>% 
  mutate(date = yday(ymd(date)),steps = as.double(steps))%>%
  group_by(interval)%>%
  mutate_at(vars(steps), ~replace_na(., mean(.,na.rm=TRUE)))
``` 

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day
```{r, echo = TRUE, message=FALSE, warning=FALSE}
tidied_data <- tidied_data%>%
  group_by(date)%>%
  summarize(total_steps_perday= sum(steps))

revised_mean <- round(mean(tidied_data$total_steps_perday, na.rm=TRUE ))
 
revised_median = round(median(tidied_data$total_steps_perday, na.rm=TRUE))

tidied_data %>%
  ggplot(aes(x=total_steps_perday)) +
  geom_histogram() + labs(title="Histogram of the tot.no of daily steps after na cleaning") +
  geom_vline(aes(xintercept = revised_mean ),linetype=2,color="blue") + 
  annotate("text", x=revised_mean-500, y=1.5, 
           label=paste("mean=",revised_mean), angle=90, size=5, color="blue",) +
  geom_vline(aes(xintercept = revised_median +40),linetype=2,color="red") + 
  annotate("text", x=revised_median+500, y=1.5, 
           label=paste("median=",revised_median), angle=90, size=5, color="red",)

```

# Are there differences in activity patterns between weekdays and weekends?



```{r, echo = TRUE, message=FALSE, warning=FALSE}
my_data <- read.csv("activity.csv")
my_data <- my_data %>% 
  mutate(date = ymd(date))

my_data <- my_data %>%
          
         group_by(date) %>%
          
          summarise(total_steps = sum(steps))




         #Create a new factor variable in the dataset with two levels – “weekday” and                                                                    “weekend” indicating whether a given date is a weekday or weekend day.


data <- read.csv("activity.csv")

tidied_data <- data %>% 
  mutate(date = ymd(date),steps = as.double(steps))%>%
  group_by(interval)%>%
  mutate_at(vars(steps), ~replace_na(., mean(.,na.rm=TRUE)))%>%
  ungroup()%>%
  group_by(date)%>%
  mutate(weekend =(wday(date)<2 | wday(date)>5))


tidied_data%>%
  group_by(weekend, interval)%>%
  summarise(interval_means = mean(steps, na.rm=TRUE))%>%
  ggplot(aes(x=interval, y= interval_means))+ geom_line() + facet_wrap(~weekend, nrow=2) + 
  ylab("average number of total steps per day") + labs(title="The distribution of average steps across a day")

```


